---
title: "NGO analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
runtime: shiny
---

```{r setup, include=FALSE}
# Loading packages using pacman to ensure they're all installed and avoiding
# catastrophic dashboard errors if they're not available
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools, purrr
)
# Using the fontawesome package for more up-to-date (and currently supported)
# Fontawesome icons
htmltools::tagList(fontawesome::fa_html_dependency())

# source_code: embed OR Github link
```


Sidebar {.sidebar}
=======================================================================

### File upload

Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.

Then upload an Excel file in the set format. <!--# Need to insert link when repository is public -->

Click 'Analyse' to analyse the data.

```{r sidebar}
# HTML upload
fileInput("html", label = "Upload html file", accept = ".html")

# Year inputs
numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel upload
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Analyse button
actionButton("click", "Analyse")
```

If desired, a copy of the analysis in PDF format may then be downloaded.

```{r sidebar_button}
# download button - using uiOutput() because of an apparent bug
# button and report are created in the second reactive chunk
uiOutput("downloadUI")
```

Dashboard
=======================================================================

Row 
-----------------------------------------------------------------------

### KPI Fundraising
```{r}
valueBoxOutput("KPI_box_fundraising")
```

### KPI OPX
```{r}
valueBoxOutput("KPI_box_OPX")
```

### KPI Private growth
```{r}
valueBoxOutput("KPI_box_private")
```

### KPI Public growth
```{r}
valueBoxOutput("KPI_box_public")
```

### KPI Private/public proportion
```{r}
valueBoxOutput("KPI_box_proportion")
```


Row
-----------------------------------------------------------------------

### Net income
```{r}
plotlyOutput("arsresultat_plotly")
```

### Total equity/liabilities
```{r}
plotlyOutput("balance_plotly")
```

Row 
-----------------------------------------------------------------------

### Revenues
```{r}
plotlyOutput("revenue_plotly")
```

### Expenses
```{r}
plotlyOutput("expenses_plotly")
```

Row 
-----------------------------------------------------------------------

### Transparency: CEO compensation
```{r}
plotlyOutput("CEO_plotly")
```

Details
=======================================================================

Row {data-height=60}
-----------------------------------------------------------------------

### NGO analysed  

```{r}
# Using htmlOutput() to use html styling
htmlOutput("title")
```

Row 
-----------------------------------------------------------------------

### Key Performance Indicators

The KPIs are based on the following calculations, using data derived from the Excel sheet:

- **Fundraising ROI**:
  - A basic measure of the return on investment of direct fundraising expenses: 
  
  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$

- **Liquid assets/obligations**:
  - A measure of the relationship between the organisation's liquid assets and its obligations:
  
  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$

- **Private donations growth**:
  - A calculation of the percentage change in private contributions based on the two most recent years:
  
  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$

- **Public grants growth**:
  - A calculation of the percentage change in public grants received based on the two most recent years:
  
$$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$

- **Private/public proportion**:
  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
  
$$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$

Row 
-----------------------------------------------------------------------

### Graphs

The dashboard includes five graphs, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

```{r reactive_title_value_boxes, include=FALSE}
# Shiny actions begin here =====================================================
# First reactive chunk =========================================================

# wait for button click
observe({
  
  # Excel data frame ===========================================================

  ## Validating and creating long version of Excel file, if possible------------
  excel_test <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))

    widen_excel_test(input$excel$datapath)
  })

  ## Creating the primary Excel data frame--------------------------------------
  exceldata <- reactive({
    widen_excel(excel_test())
  })

  # Value Boxes ================================================================

  # All boxes use the colour "primary" and the icon "fa-solid fa-circle-xmark"
  # if the value is NA

  ## Fundraising ROI------------------------------------------------------------
  # Creating Fundraising ROI value
  KPI_fundraising_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        transmute(
          # Calculating the value as a percentage; removing all other
          KPI_fundraising_num =
            (sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) - expenses_fundraising) /
              expenses_fundraising * 100
        ) |>
        # just want the single number
        pull() |>
        first()
    }
  })

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  output$KPI_box_fundraising <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_fundraising_num()),
      caption = "Fundraising ROI",
      # Colour depends on value
      color = if (is.na(KPI_fundraising_num())) {
        "primary"
      } else if (KPI_fundraising_num() < 100) {
        "danger"
      } else if (KPI_fundraising_num() < 200) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_fundraising_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-donate"
      }
    )
  })

  ## OPX KPI--------------------------------------------------------------------
  # Creating OPX KPI
  KPI_OPX_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        # Calculating the value as a %
        transmute(
          KPI_OPX =
            sum(revenue_private_contributions,
              revenue_investments, assets_cash_in_bank_and_cash_equivalents,
              na.rm = TRUE
            ) /
              sum(liability_short_term_grants_payable,
                liability_revocable_endowments,
                liability_long_term_grants_payable,
                na.rm = TRUE
              ) * 100
        ) |>
        # Just the single number
        pull() |>
        first()
    }
  })

  # Creating value box for OPX KPI
  output$KPI_box_OPX <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_OPX_num()),
      caption = "Liquid assets / obligations",
      # colour depends on value
      color = if (is.na(KPI_OPX_num())) {
        "primary"
      } else if (KPI_OPX_num() < 100) {
        "warning"
      } else {
        "success"
      },
      # icon depends on value
      icon = if (is.na(KPI_OPX_num())) {
        "fa-solid fa-circle-xmark"
      } else if (KPI_OPX_num() < 100) {
        "fa-solid fa-scale-unbalanced"
      } else {
        "fa-solid fa-scale-balanced"
      }
    )
  })

  ## Private donations growth---------------------------------------------------
  # Creating private donations growth
  KPI_private_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth =
            (revenue_private_contributions[1] -
              revenue_private_contributions[2]) /
              revenue_private_contributions[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for private donations growth
  output$KPI_box_private <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_private_num()),
      caption = "Private donation growth",
      # colour depends on value
      color = if (is.na(KPI_private_num())) {
        "primary"
      } else if (KPI_private_num() < 0) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_private_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fas fa-hands-helping"
      }
    )
  })

  ## Public grants growth-------------------------------------------------------
  # Creating public grants growth
  KPI_public_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth =
            (revenue_public_grants[1] - revenue_public_grants[2]) /
              revenue_public_grants[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for public grants growth
  output$KPI_box_public <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_public_num()),
      caption = "Public grants growth",
      # colour depends on value
      color = if (is.na(KPI_public_num())) {
        "primary"
      } else if (KPI_public_num() < 0) {
        "warning"
      } else {
        "success"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_public_num())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-building-columns"
      }
    )
  })

  ## Public/private proportion--------------------------------------------------
  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # calculating the value as a %
        transmute(
          proportion =
            sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) /
              sum(revenue_public_grants, revenue_private_contributions,
                revenue_membership_fees,
                na.rm = TRUE
              ) * 100
        ) |>
        pull() |>
        # getting the most recent year
        first()
    }
  })

  # Creating value box for private/public proportion
  output$KPI_box_proportion <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_proportion()),
      caption = "Private donations",
      # Adding in the same check as elsewhere - colours match if all are NA
      color = if (is.na(KPI_proportion())) {
        "primary"
      } else {
        "info"
      },
      # icon depends on whether or not the value is NA
      icon = if (is.na(KPI_proportion())) {
        "fa-solid fa-circle-xmark"
      } else {
        "fa-solid fa-chart-pie"
      }
    )
  })

  ## Excel data test -----------------------------------------------------------
  # (also below, in the second chunk)
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
  # the button click
}) |> bindEvent(input$click)
```


```{r reactive_report_plots, include=FALSE}
# Beginning of the second reactive chunk =======================================
# It's necessary to split the observers into separate parts, with all necessary
# code duplicated, because of R's limitations.

observe({
  
  # Data frames ================================================================
  ## Creating data frame from html file-----------------------------------------
  html_currency <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))

    # Testing and creating data frame
    html_test_create(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  htmldata <- reactive({
    read_proff_two(html_currency())
  })


  # Recreating, to make available to this chunk as well
  ## Validating and creating long version of Excel file, if possible------------
  excel_test <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))

    widen_excel_test(input$excel$datapath)
  })

  ## Creating the primary Excel data frame--------------------------------------
  exceldata <- reactive({
    widen_excel(excel_test())
  })
  
  # NGO title ==================================================================
  NGO <- reactive({
    # checking for the required column again
    if ("firm" %in% names(excel_test())) {
      # fetching the name
      excel_test() |>
        # Turning the needed column into a list
        pull(firm) |>
        # discarding all NAs in case the name is somewhere
        purrr::discard(is.na) |>
        # choosing the first available name
        first()
    } else {
      validate("Please upload data using the specified Excel template.")
    }
  })

  title <- reactive({
    # testing if NGO() is NA
    if (is.na(NGO())) {
      # Printing warning if NA
      "Name of NGO missing"
      # Printing name of NGO
    } else {
      NGO()
    }
  })
  
  # Creating NGO title using html formatting (colour based on MetBrewer
  # "Hokusai2")
  output$title <- renderText({
    paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
  })
  
  # Report Output ==============================================================

  ## Download button -----------------------------------------------------------
  output$downloadUI <- renderUI({
    if (available_excel() && available_html()) {
      downloadButton("report", "Download report")
    }
  })
  
  ## Data exists condition testing----------------------------------------------
  # Essentially pretty random tests to connect to the individual
  # validate() functions upstream in the reactive chain
  
  # Excel
  available_excel <- reactive({
    TRUE
  }) |>
    bindEvent(setequal(
      names(exceldata()),
      c("year", "firm", "category", "sub_category", "name", "value")
    ))

  # HTML
  available_html <- reactive({
    TRUE
  }) |> bindEvent("lederlonn" %in% names(htmldata()))
  
  ## Report download and parameters --------------------------------------------
  output$report <- downloadHandler(

    # creating the file name using the NGO title and system date
    filename = function() {
      str_c(title(), "_", Sys.Date(), ".pdf")
    },
    content = function(file) {
      # Based on code in Mastering Shiny -- creating a temporary directory and
      # file
      tempReport <- file.path(tempdir(), "report.Rmd")
      file.copy("report.Rmd", tempReport, overwrite = TRUE)

      # parameters passed to the R Markdown document
      params <- list(
        htmlfile = input$html$datapath,
        excelfile = input$excel$datapath,
        start = input$start,
        end = input$end
      )

      # notification -- disappears when report generation is complete
      id <- showNotification(
        "Generating report...",
        duration = NULL,
        closeButton = FALSE
      )
      on.exit(removeNotification(id), add = TRUE)

      # R Markdown render
      rmarkdown::render(tempReport,
        output_file = file,
        params = params,
        envir = new.env(parent = globalenv())
      )
    }
  )

  # Plots ======================================================================

  ## Annual result--------------------------------------------------------------
  arsresultat_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # renaming for plotly hover
        rename(
          Year = year,
          `Net income` = arsresultat,
          `Net positive` = annual_profit
        ) |>
        ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$annual_profit) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Net income (in thousands)"
        )
    }
  })

  # Creating plotly output
  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot()) |>
      # ggplot subtitles and captions are lost when using ggplotly(), a
      # currency warning caption is therefore added in plotly, if needed, using
      # a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  ## Balance--------------------------------------------------------------------
  balance_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # rename for plotly hover
        rename(
          Year = year,
          `Total equity/liabilities` = sum_egenkapital_og_gjeld,
          `Positive equity` = balance_positive
        ) |>
        ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$balance_positive) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Total equity/liabilities (in thousands)",
        )
    }
  })

  # Creating plotly output
  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  ## Expenses-------------------------------------------------------------------
  expenses_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else if (excel_zero_check("Expenses") == 0) {
      zero_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          Other = expenses_other,
          Administrative = expenses_administrative,
          Fundraising = expenses_fundraising,
          `Program services` = expenses_program_services
        ) |>
        # Pivoting to use fill
        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
        # Ordering for the plot
        mutate(
          Expenses = ordered(Expenses,
            levels =
              c("Program services", "Other", "Fundraising", "Administrative")
          )
        ) |>
        ggplot(aes(Year, value, fill = Expenses)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        scale_fill_met_d("Hokusai2") +
        labs(x = "", y = "", title = "Expenses (in thousands)")
    }
  })

  # Creating plotly output
  output$expenses_plotly <- renderPlotly(
    ggplotly(expenses_plot())
  )

  ## Revenue--------------------------------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else if (excel_zero_check("Revenue") == 0) {
      zero_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        # Ordering for the plot
        mutate(
          `Revenue source` =
            ordered(`Revenue source`,
              levels =
                c(
                  "Other", "Commercial activities",
                  "Operational activities", "Investments",
                  "Membership fees", "Private contributions",
                  "Public grants"
                )
            )
        ) |>
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

  # Creating plotly output
  output$revenue_plotly <- renderPlotly(
    ggplotly(revenue_plot())
  )

  ## CEO compensation-----------------------------------------------------------
  CEO_plot <- reactive({
    # Testing for the presence of data
    if (sum(!is.na(htmldata()$lederlonn)) == 0 &&
      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      # Creating plot if any data exist
      # Note that the plot will still look empty, but have a legend and scales,
      # if all values are explicit zeroes
      htmldata() |>
        # preparing for pivoting
        select(
          year, lederlonn, leder_annen_godtgjorelse
        ) |>
        # Compensation in ordinary numbers
        mutate(
          lederlonn = lederlonn * 1000,
          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
        ) |>
        # renaming for the purposes of the plot
        rename(
          Year = year,
          Salary = lederlonn,
          `Other compensation` = leder_annen_godtgjorelse
        ) |>
        # pivoting to join the two values to use fill
        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
        ggplot(aes(Year, value, fill = Compensation)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # pretty colours
        scale_fill_met_d("Hokusai2", direction = -1) +
        labs(
          x = "", y = "Total compensation", title = "CEO compensation"
        )
    }
  })

  # Creating plotly output
  output$CEO_plotly <- renderPlotly(
    ggplotly(CEO_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  # Reactive functions =========================================================

  ## Currency scale-------------------------------------------------------------
  # Reactive function to check for currency and use currency code or ?
  currency_scale <- reactive({
    scale_y_continuous(
      # Checking for uniform currency code
      labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
        # Using the currency code if uniform
        label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
        # alternative prefix if multiple currencies in the data
        label_dollar(prefix = "? ")
      )
    )
  })

  ## Plotly caption-------------------------------------------------------------
  # Reactive function to create plotly caption (ggplot captions and subtitles
  # are lost in the conversion) (shamelessly stolen from the Internet, with
  # minor modifications).
  # Checks for uniform currency code and prints a warning
  # if there are multiple currency codes.
  plotly_caption <- reactive({
    list(
      x = 1, y = -0.1,
      text = ifelse(length(unique(html_currency()$valutakode)) == 1,
        "", "Note: not all values are denominated in the same currency"
      ),
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 11)
    )
  })

  ## scale maximum--------------------------------------------------------------
  # Reactive function to create the maximum value for revenue and expenses plots
  maximum <- reactive({
      excel_test() |>
        # removing the firm and sub-category columns
        select(-firm, -sub_category) |>
        # filtering to retain only revenue and expenses entries
        filter(category == "Revenue" | category == "Expenses") |>
        # grouping and summarising to calculate maximum value
        group_by(year, category) |>
        summarise(
          sum = sum(value, na.rm = TRUE)
        ) |>
        # Creating list of sums
        pull() |>
        # selecting the highest value
        max()
  })

  ## Excel data test (also above)-----------------------------------------------
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # Testing if there are any characters instead of numbers (data probably
    # hopelessly corrupted)
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
  
  ## Excel zero test------------------------------------------------------------
  excel_zero_check <- function(type) {
    excel_test() |>
      # Choosing the specified category (i.e. Revenue or Expenses)
      filter(
        category == as.character(all_of(type))
      ) |>
      # Calculating if there are any zeroes
      summarise(
        zero = sum(value != 0, na.rm = TRUE)
      ) |>
      # just the single number
      pull()
  }
  
  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
}) |> bindEvent(input$click)
```

```{r non-reactive functions, include=FALSE}
# Data reading functions =======================================================

## Functions to read html file -------------------------------------------------
html_test_create <- function(file, start, end) {
  # Reading file and node
  htmltest <- read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")

  ### A series of tests to catch problems before they spit out errors-----------
  # Checking if node exists
  if (class(htmltest) == "xml_node") {
    htmltable <- htmltest |>
      html_table()
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Basic check if the table is correct
  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
    htmlready <- htmltable
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Testing if the correct years have been entered
  if (as.character(all_of(start)) %in% names(htmlready) &&
    as.character(all_of(end)) %in% names(htmlready)) {
    # Creating the first part of the data frame
    read_proff_one(htmltable, start, end)
  } else {
    validate("Please ensure you have entered the correct start and end years.")
  }
}

### First html transformation function, invoked by html_test_create() ----------
# (stopping to allow checks for differing currency codes)
read_proff_one <- function(object, start, end) {
  object |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column to avoid having to deal with NAs later
    fill(
      valutakode,
      .direction = "downup"
    )
}

### Second html transformation function, creating the data frame ---------------
read_proff_two <- function(object) {
  object |>
    # removing superfluous currency column
    select(-valutakode) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots (red/green profit/loss)
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}

## Functions to read Excel file ------------------------------------------------
### First Excel transformation -------------------------------------------------
# creating df used by NGO title function, maximum scale function, and excel zero
# check
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |>
    # cleaning names to enable later operations
    clean_names()
}

### Second Excel function - creating primary Excel data frame-------------------
widen_excel <- function(object) {
  # testing if all columns are present
  if (setequal(
    names(object),
    c("year", "firm", "category", "sub_category", "name", "value")
  )) {
    object |>
      # uniting category and name to deal with multiple 'Other'
      unite(
        entry, c(category, name)
      ) |>
      # deleting Sub-category and Firm
      select(-sub_category, -firm) |>
      # pivoting to create 'tidy' variables
      pivot_wider(names_from = entry, values_from = value) |>
      # cleaning names
      clean_names()
    # Notification if all columns are not present
  } else {
    validate("Please upload data using the specified Excel template.")
  }
}

# Support functions ============================================================

## KPI_to_string ---------------------------------------------------------------
# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
  # avoid passing an NA to subsequent operations.)
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}

## scale_fill_green ------------------------------------------------------------
# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}

## no_information plot ---------------------------------------------------------
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  empty_plot("No information available")
}

## bad_data plot ---------------------------------------------------------------
# Function to print a 'bad data' plot, if necessary
bad_data <- function() {
  empty_plot("Something is wrong\nwith the Excel data.")
}

## zero_data_plot --------------------------------------------------------------
zero_data <- function() {
  empty_plot(
    "All values are zero or missing.
    This is unusual and may
    indicate a data problem.")
}

## empty_plot-------------------------------------------------------------------
# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    # annotating the plot with the specified warning message
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 6, colour = met.brewer("Hokusai2", 1)
    )
}
```

