---
date: "`r Sys.Date()`"
output: 
  pdf_document:
    extra_dependencies: "subfig"
params:
  htmlfile: NA
  excelfile: NA
  start: NA
  end: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.asp = 0.618, out.width = "75%", fig.align = "center")

# Repeating loading the libraries, in case the information isn't passed on from
# the dashboard
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools, purrr
)
```

<!--# The title element of the YAML block is further down, as it relies on data from the Excel sheet. -->

```{r data}
# Data reading =================================================================

## Function to read html file --------------------------------------------------
read_proff <- function(file, start, end) {
  # Specifying encoding to deal with occasional errors
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    # filling the currency column
    fill(
      valutakode,
      .direction = "updown"
    ) |>
    # removing dates and redundant currency
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}

# Creating a data frame for currency (for some reason the functions in the
# dashboard create bad data, but *only* in a *parameterised* R Markdown report).
read_proff_currency <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill(
      valutakode,
      .direction = "downup"
    )
}

## Function to read Excel file -------------------------------------------------
widen_excel <- function(file) {
  read_excel(file) |>
    # uniting Category and Name to deal with multiple 'Other'
    unite(
      entry, c(Category, Name),
      sep = "_"
    ) |>
    # deleting Sub-category and Firm
    select(-`Sub-category`, -Firm) |>
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = Value) |>
    # cleaning names
    clean_names()
}
### Reading html ---------------------
# Creating the main data frame
htmldata <- read_proff(
  params$htmlfile,
  as.character(params$start), as.character(params$end)
)

# Creating the currency data frame
html_currency <- read_proff_currency(params$htmlfile,
  as.character(params$start), as.character(params$end))

### Reading excel ---------------------
exceldata <- widen_excel(params$excelfile)

# NGO title ==================================================================

NGO <- read_excel(params$excelfile) |>
  # Turning the needed column into a list
  pull(Firm) |>
  # discarding all NAs in case the name is somewhere
  purrr::discard(is.na) |>
  # choosing the first available name
  first()

title <- # testing if NGO() is NA
  if (is.na(NGO)) {
    # Printing warning if NA
    "Name of NGO missing"
    # Printing name of NGO
  } else {
    NGO
  }
```

<!--# Report title-->

---
title: "NGO Analysis: `r title`"
---

```{r data_warnings}
# Excel data test =============================================================
# Function to test for Excel data quality (basic)
datatest <-
  # let me know if anyone has a better idea for testing
  if (mean(unlist(lapply(exceldata, is.character))) != 0) {
    TRUE
  } else {
    FALSE
  }

## Creating the Excel warning text --------------------------------------------
excel_warning <- if (datatest == TRUE) {
  "***    Warning: something is wrong with the Excel data.***"
} else {
  ""
}

# HTML warnings ===============================================================
# Creating the arsresultat warning text
arsresultat_plot_warning <- if (sum(!is.na(htmldata$arsresultat)) == 0) {
  "*    No information about net income/loss is available.*"
} else {
  ""
}

# Creating the balance warning text
balance_plot_warning <- if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
  "*    No information about total equity/liabilities is available.*"
} else {
  ""
}

# Creating the CEO_plot warning text
CEO_plot_warning <- if (sum(!is.na(htmldata$lederlonn)) == 0 &&
  sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
  "*    No information about CEO compensation is available.*"
} else {
  ""
}

# Creating general warning
general_warning <- if (datatest == TRUE ||
  sum(!is.na(htmldata$arsresultat)) == 0 || 
  sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0 ||
  (sum(!is.na(htmldata$lederlonn)) == 0 &&
    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0)) {
  "## Note:"
} else {
  ""
}
```

<!--# Functions that don't rely on any of the parameters passed from the dashboard -->

```{r static_functions}
## KPI_to_string ---------------------------------------------------------------
# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable) {
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting. (The check for NA is needed to
  # avoid passing an NA to subsequent operations.)
  if (is.na(variable)) {
    NA
  } else if (variable == Inf) {
    NA
  } else {
    # Shortening (lengthening) the value, adding %; making the text bold
    str_c("**", format(variable, digits = 3), "%**")
  }
}

## scale_fill_green ------------------------------------------------------------
# Function with slightly nicer colours than default; ensuring uniformity; +
# ensuring positive values are always green through a test for the presence of
# FALSE values
scale_fill_green <- function(variable) {
  scale_fill_met_d("Tsimshian",
    direction = ifelse(
      mean(variable,
        na.rm = TRUE
      ) == 1, -1, 1
    )
  )
}
```

<!--# Functions that rely on parameters from the dashboard -->

```{r functions_reliant}
## scale maximum--------------------------------------------------------------
# Function to create the maximum value for revenue and expenses plots
maximum <- if (datatest == TRUE) {
  NA
} else {
  # reading the Excel sheet anew
  read_excel(params$excelfile) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
}

## Currency scale-------------------------------------------------------------
# Function to check for currency and use currency code or ?
currency_scale <-
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(html_currency$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(unique(html_currency$valutakode), " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )

## Currency caption-----------------------------------------------------------
currency_caption <- ifelse(length(unique(html_currency$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency."
    )
```

<!--# Creating KPIs -->

```{r KPIs}
## Fundraising ROI------------------------------------------------------------
# Creating Fundraising ROI value
# testing for data quality, returning NA if bad
fundraising <- if (datatest == TRUE) {
  NA
} else {
  {
    exceldata |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  } |> KPI_to_string()
}

## OPX KPI--------------------------------------------------------------------
# Creating OPX KPI
KPI_OPX <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      # most recent year
      filter(year == max(year)) |>
      # Calculating the value as a %
      transmute(
        KPI_OPX =
          sum(revenue_private_contributions,
            revenue_investments, assets_cash_in_bank_and_cash_equivalents,
            na.rm = TRUE
          ) /
            sum(liability_short_term_grants_payable,
              liability_revocable_endowments,
              liability_long_term_grants_payable,
              na.rm = TRUE
            ) * 100
      ) |>
      # Just the single number
      pull() |>
      first() |>
      KPI_to_string()
  }

## Private donations ---------------------------------------------------------
# Creating private donations growth
KPI_private <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_private_contributions[1] -
            revenue_private_contributions[2]) /
            revenue_private_contributions[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first() |>
      KPI_to_string()
  }

## Public grants growth-------------------------------------------------------
# Creating public grants growth
KPI_public <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      # Calculating the two most recent years as a %
      transmute(
        growth =
          (revenue_public_grants[1] - revenue_public_grants[2]) /
            revenue_public_grants[2] * 100
      ) |>
      # Pulling the first instance of the number
      pull() |>
      first() |>
      KPI_to_string()
  }

## Public/private proportion--------------------------------------------------
# Creating public/private proportion
KPI_proportion <-
  # testing for data quality, returning NA if bad
  if (datatest == TRUE) {
    NA
  } else {
    exceldata |>
      transmute(
        proportion =
          sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) /
            sum(revenue_public_grants, revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) * 100
      ) |>
      pull() |>
      first() |>
      KPI_to_string()
  }

### NA warning-----------------------------------------------------------------
# Testing for the presence of NAs; printing warning if there are any
KPI_warning <- if (is.na(fundraising) || is.na(KPI_OPX) || is.na( KPI_private) ||
  is.na(KPI_public) || is.na(KPI_proportion)) {
  "*A value of 'NA' indicates that there is something wrong with the Excel data, that the necessary data are not available, or that the data are unsuited to an automatic calculation of the KPI.*"
} else {
  ""
}
```

# Key Performance Indicators

The KPIs are based on the following calculations, using data derived from the Excel sheet:

- **Fundraising ROI:** `r fundraising`
  - A basic measure of the return on investment of direct fundraising expenses:
  $$\frac{\text{private contributions} + \text{membership fees} - \text{fundraising expenses}}{\text{fundraising expenses}} \times 100$$

- **Liquid assets/obligations:** `r KPI_OPX`
  - A measure of the relationship between the organisation's liquid assets and its obligations:
  $$\frac{\text{private contributions} + \text{investment income} + \text{cash in bank and cash equivalents}}{\text{short term grants payable} + \text{revocable endowments} + \text{long term grants payable}} \times 100$$

- **Private donations growth:** `r KPI_private`
  - A calculation of the percentage change in private contributions based on the two most recent years:
  $$\frac{\text{private contributions}_2 - \text{private contributions}_1}{\text{private contributions}_1}\times 100$$

- **Public grants growth:** `r KPI_public`
  - A calculation of the percentage change in public grants received based on the two most recent years:
  $$\frac{\text{public grants}_2 - \text{public grants}_1}{\text{public grants}_1} \times 100$$

- **Private/public proportion:** `r KPI_proportion`
  - Calculates the percentage of non-commercial or activity-based income that is derived from private contributions and membership fees as opposed to public grants:
  $$\frac{\text{private contributions} + \text{membership fees}}{\text{public grants} + \text{private contributions} + \text{membership fees}} \times 100$$

`r KPI_warning`

<!--# Latex command to force new page - ensures the plot heading doesn't move up to the first page-->

\newpage

# Graphs

<!--# Warning header, if necessary -->

`r general_warning`

<!--# specific alerts -->

`r excel_warning`

`r arsresultat_plot_warning`

`r balance_plot_warning`

`r CEO_plot_warning`

<!--# Creating plots -->

<!-- But of course a bit of raw LaTeX is necessary to print a left quotation mark -->

```{r arsresultat_plot, fig.cap="Based on the entry \`Årsresultat' in the Proff.no data."}
## Annual result--------------------------------------------------------------
arsresultat_plot <-
  # testing for the presence of data
  if (sum(!is.na(htmldata$arsresultat)) == 0) {
    NA
  } else {
    htmldata |>
      # renaming for plotly hover
      rename(
        Year = year,
        `Net income` = arsresultat,
        `Net positive` = annual_profit
      ) |>
      ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata$annual_profit) +
      labs(
        x = "", y = "", caption = currency_caption, title = "Net income (in thousands)"
      ) +
      theme_light() +
      theme(legend.position = "none")
  }
# Printing plot if the data are available
if (sum(!is.na(htmldata$arsresultat)) != 0) {
  arsresultat_plot
}
```

```{r balance_plot, fig.cap="Based on the entry \`Sum egenkapital og gjeld' in the Proff.no data."}
## Balance--------------------------------------------------------------------
balance_plot <-
  # testing for the presence of data
  if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) == 0) {
    # Printing 'no information' plot if no data
    NA
  } else {
    htmldata |>
      # rename for plotly hover
      rename(
        Year = year,
        `Total equity/liabilities` = sum_egenkapital_og_gjeld,
        `Positive equity` = balance_positive
      ) |>
      ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # using a function to ensure uniform colours irrespective of data
      scale_fill_green(htmldata$balance_positive) +
      labs(
        x = "", y = "",
        caption = currency_caption, title = "Total equity/liabilities (in thousands)"
      ) +
      theme_light() +
      theme(legend.position = "none")
  }

# Printing the plot if the data are available
if (sum(!is.na(htmldata$sum_egenkapital_og_gjeld)) != 0) {
  balance_plot
}
```


```{r revenue_expenses_plot, fig.asp=1, out.width="50%", fig.cap="Revenue and Expenses", fig.subcap=c("Based on the category \`Revenue' in the Excel sheet.", "Based on the category \`Expenses' in the Excel sheet.")}
## Revenue--------------------------------------------------------------------
revenue_plot <-
  # testing for data quality
  if (datatest == TRUE) {
    # printing 'bad data' plot if data are bad
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        `Private contributions` = revenue_private_contributions,
        `Membership fees` = revenue_membership_fees,
        `Public grants` = revenue_public_grants,
        Investments = revenue_investments,
        `Operational activities` = revenue_operational_activities,
        `Commercial activities` = revenue_commercial_activities,
        Other = revenue_other
      ) |>
      # Pivoting to use fill
      pivot_longer(
        c(`Private contributions`:Other),
        names_to = "Revenue source"
      ) |>
      # Ordering for the plot
      mutate(
        `Revenue source` =
          ordered(`Revenue source`,
            levels =
              c(
                "Other", "Commercial activities",
                "Operational activities", "Investments",
                "Membership fees", "Private contributions",
                "Public grants"
              )
          )
      ) |>
      ggplot(aes(Year, value, fill = `Revenue source`)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      # Specifying the number of values, as the base palette only includes 6
      scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
      labs(x = "", y = "", title = "Revenue (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }

# Printing the plot if the data are OK
if (datatest == FALSE) {
  revenue_plot
}

## Expenses -------------------------------------------------------------------
expenses_plot <-
  # testing for data quality
  if (datatest == TRUE) {
    # printing 'bad data' plot if data are bad
    NA
  } else {
    exceldata |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      # Ordering for the plot
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c("Program services", "Other", "Fundraising", "Administrative")
        )
      ) |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      scale_y_continuous(
        # Excel data are assumed to be in NOK; if this changes, code based on
        # the currency_code() function may be adapted
        labels = label_dollar(prefix = "NOK "),
        # Basing lower limit on the data (not suppressing erroneous negative
        # numbers), and the upper limit on the maximum value in revenues and
        # expenses, to match the two plots
        limits = c(NA, maximum)
      ) +
      scale_fill_met_d("Hokusai2") +
      labs(x = "", y = "", title = "Expenses (in thousands)") +
      theme_light() +
      theme(title = element_text(size = 14))
  }

# Printing the plot if the data are OK
if (datatest == FALSE) {
  expenses_plot
}
```

```{r compensation_plot, out.width="100%", fig.cap="Based on the category \`Lederlønn' in the Proff.no data. Not in thousands. Includes the entries \`Lønn' and \`Leder annen godtgjørelse'. A common measure of transparency, but should normally be evaluated in context."}
## CEO compensation-----------------------------------------------------------
CEO_plot <-
  # Testing for the presence of data
  if (sum(!is.na(htmldata$lederlonn)) == 0 &&
    sum(!is.na(htmldata$leder_annen_godtgjorelse)) == 0) {
    # Returning NA if no data
    NA
  } else {
    # Creating plot if any data exist
    # Note that the plot will still look empty, but have a legend and scales,
    # if all values are explicit zeroes
    htmldata |>
      # preparing for pivoting
      select(
        year, lederlonn, leder_annen_godtgjorelse
      ) |>
      # Compensation in ordinary numbers
      mutate(
        lederlonn = lederlonn * 1000,
        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
      ) |>
      # renaming for the purposes of the plot
      rename(
        Year = year,
        Salary = lederlonn,
        `Other compensation` = leder_annen_godtgjorelse
      ) |>
      # pivoting to join the two values to use fill
      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
      ggplot(aes(Year, value, fill = Compensation)) +
      geom_col() +
      # using currency scale function to add correct currency code or ?
      currency_scale +
      # pretty colours
      scale_fill_met_d("Hokusai2", direction = -1) +
      labs(
        x = "", y = "", title = "CEO compensation"
      ) +
      theme_light() +
      theme(legend.position = "bottom")
  }

# Printing the plot if the data are available
if (sum(!is.na(htmldata$lederlonn)) != 0 ||
  sum(!is.na(htmldata$leder_annen_godtgjorelse)) != 0) {
  CEO_plot
}
```
